---
# title: "Dineout"
author: "Dillon"
date: "2023-11-20"
output: pdf_document
---

# Setup

```{r setup, echo=TRUE}
# libraries
library(sf)
library(dplyr)
library(ggplot2)
library(jpeg)

# data sets
restaurants = read.csv("restaurants.csv")
movements = read.csv("movements.csv")

# function for calculating buffer based on an input in sq. footage (circle->radius in feet->meters->coords)
calc_buffer <- function(size)
{
  return((((size / pi)^(0.5))*0.3048)/111111)
}
```

# Cleaning up movements

```{r, echo=TRUE}
# Sort by datetime
movements <- movements %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  arrange(datetime)

# Filter duplicates
movements <- movements %>%
  mutate(err = if_else(datetime > lag(datetime, default = as.POSIXct("2019-12-31 23:59:59", format = "%Y-%m-%d %H:%M:%S")), 0, 1)) %>%
  filter(err == 0)
```

# EDA

```{r, echo=TRUE}
# Categories of restaurants
rest_types <- restaurants %>%
  count(Category)
rest_types

restaurants

# Time range
#print(first(movements$datetime))
#print(last(movements$datetime))
#print(difftime(last(movements$datetime), first(movements$datetime)))
```

# Parameters

```{r, echo=TRUE}
# buffers for categories of restaurants (converted from ft. sq.)
buffer_full = calc_buffer(10000)+0.0001
buffer_limited = calc_buffer(3740)+0.0001
buffer_market = calc_buffer(22500)+0.0001
buffer_supermarc = calc_buffer(48415)+0.0001
buffer_warehouse = calc_buffer(140000)+0.0001
buffer_conv = calc_buffer(2500)+0.0001

# threshold for considering a new visit (in minutes)
time_gap_threshold <- 6
```

# Convert restaurant set into sf object and add buffers (dependent on category)

```{r, echo=TRUE}
# Convert restaurants to sf object. Keep "Name" and "Category" variables. Use Long and Lat for coordinates.
rest_sf <- st_as_sf(restaurants[, c("Name", "Category", "Longitude", "Latitude")], coords = c("Longitude", "Latitude"))

# Add column for appropriate buffers then apply them to the object
rest_sf <- rest_sf %>% 
  mutate(buffer = case_when(
    Category == "Full-Service Restaurants" ~ buffer_full,
    Category == "Limited-Service Restaurants" ~ buffer_limited,
    Category == "Fruit & Vegetable Markets" ~ buffer_market,
    Category == "Supermarkets/Other Grocery (Exc Convenience) Strs" ~ buffer_supermarc,
    Category == "Warehouse Clubs & Supercenters" ~ buffer_warehouse,
    Category == "Convenience Stores" ~ buffer_conv
  ))
rest_sf <- st_buffer(rest_sf, dist = rest_sf$buffer)

plot(select(rest_sf, -Category, -buffer), col = "red", main = "Restaurants")
```

# For each point in the movement set, find which (if any) restaurant's buffer it is within.
* Creates a data frame of TRUE and FALSE where movements are observations and each restaurant's buffer is a variable

```{r, echo=TRUE}
# Use long and lat in movements data set against restaurant buffers using st_within
movements_within <- as.data.frame(st_within(st_as_sf(movements, coords = c("longitude", "latitude")), rest_sf, sparse = FALSE))
```

# Reduce movements_within to only contain id of restaurant within or 0 if not within any. Bind the datetime as well.

```{r, echo=TRUE}
# If whole row is FALSE, set to 0, otherwise set to index of TRUE. Convert to a data frame.
movements_within <- data.frame(location = ifelse(rowSums(movements_within) == 0, 0, max.col(movements_within)))

# Bind datetime from movements set (convert to POSIX as well)
movements_within <- cbind(movements_within, datetime = as.POSIXct(movements$datetime))

```

# View visits

```{r, echo=TRUE}
head(movements_within)
```

# Cleanup for memory 

```{r, echo=FALSE}
rm(movements)
gc()
```

# Collect consecutive time spent in a location into a single "visit"

```{r, echo=TRUE}
# Create data frame for visits
visits <- data.frame(loc = numeric(0), start = character(0), end = character(0), stringsAsFactors = FALSE)

# New visit starts when prev location is different from current then use cumsum to give each an ID
movements_within <- movements_within %>%
  mutate(new_visit = location != lag(location, default = FALSE))
movements_within <- movements_within %>%
  mutate(visit_id = cumsum(new_visit))

# Filter out non-visits and reduce the group forming a visit (by visit_id) into location, start datetime and end datetime
visits <- movements_within %>%
  filter(location != 0) %>%
  group_by(visit_id) %>%
  summarize(loc = first(location), start = first(datetime), end = last(datetime)) %>%
  ungroup() %>%
  select(-visit_id)

```

# For consecutive visits to the same location, if the time between visits in under threshold (in minutes), merge into one.
* Could be stepping out to car, or to smoke, etc. 

```{r, echo=TRUE}
# Check if location of next visit is same location And if the gap in time is under threshold.
# If yes, time_gap = difftime, otherwise make it 0. If a gap is present, mark the leading row for removal 
# and replace the end time with proper end. Then remove redundant (marked) and reduce to loc, start, end.
visits <- visits %>%
  mutate (time_gap = ifelse(loc == lead(loc) & difftime(lead(start), end, units = "mins") <= time_gap_threshold, 
                      difftime(lead(start), end, units = "mins"), 0),
          rm = ifelse(lag(time_gap, default = 0) > 0, 1, 0),
          end = if_else(time_gap > 0, as.POSIXct(lead(end)), end)
         ) %>%
  filter(rm == 0) %>%
  select(-rm)
```

# Visualize visits

```{r, echo=TRUE}

# Column for visit length
visits <- mutate(visits, length = round(difftime(end, start, units = "mins"), 2))

# Replace loc with restaurant names and add categories from the data set
visits$Category <- restaurants$Category[visits$loc] 
visits$loc <- restaurants$Name[visits$loc]

# Filter out visits under category dependent length
visits <- visits %>%
  filter(case_when(
    Category == "Full-Service Restaurants" ~ length > 15,
    Category == "Limited-Service Restaurants" ~ length > 4,
    Category == "Fruit & Vegetable Markets" ~ length > 5,
    Category == "Supermarkets/Other Grocery (Exc Convenience) Strs" ~ length > 5,
    Category == "Warehouse Clubs & Supercenters" ~ length > 5,
    Category == "Convenience Stores" ~ length > 3
  ))
```

```{r, echo=TRUE}
# Tabular of min max mean stdev of visit by location, category, etc.
print("Visit Stats By Category")
category_stats <- visits %>%
  group_by(Category) %>%
  summarize(
    min = min(length),
    max = max(length),
    avg = mean(length),
    sdev = sd(length),
    total = n()
  )
print(category_stats)

# Tabular of min max mean stdev of visit by location, category, etc.
print("Visit Stats By Location")
loc_stats <- visits %>%
  group_by(loc) %>%
  summarize(
    min = min(length),
    max = max(length),
    avg = mean(length),
    sdev = sd(length),
    total = n()
  )
print(loc_stats)

# Bar plots for visits
cate_bar <- ggplot(category_stats, aes(x = Category, y = total)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_text(aes(label = total), vjust = -0.5, color = "black") +
  labs(title = "Visits by Category", x = "", y = "Visits") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
        plot.title = element_text(hjust = 0.5),
    plot.margin = margin(t = 20, r = 20, b = 1, l = 20, unit = "pt"))+
  ylim(0, max(category_stats$total) + 1000) 
loc_bar <- ggplot(loc_stats, aes(x = loc, y = total)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  geom_text(aes(label = total), angle = 33, vjust = 0, hjust = 0, color = "black") +
  labs(title = "Visits by Location", x = "", y = "Visits") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
        plot.title = element_text(hjust = 0.5),
    plot.margin = margin(t = 20, r = 20, b = 1, l = 20, unit = "pt"))+
  ylim(0, max(loc_stats$total) + 500)+
  scale_x_discrete(expand = c(0.05, 0))

# Boxplot by Category
cate_box <- ggplot(visits, aes(x = Category, y = length)) +
  geom_boxplot() +
  labs(title = "Visit Length by Category", x="", y = "Visit Length") +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 8),
        plot.title = element_text(hjust = 0.5),
    plot.margin = margin(t = 20, r = 20, b = 1, l = 20, unit = "pt"))

# Boxplot by Location
loc_box <- ggplot(visits, aes(x = loc, y = length)) +
  geom_boxplot() +
  labs(title = "Visit Length by Location", x="", y = "Visit Length") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
        plot.title = element_text(hjust = 0.5),
    plot.margin = margin(t = 20, r = 20, b = 1, l = 20, unit = "pt"))

cate_bar
loc_bar
cate_box
loc_box

# Time between visits
visits <- visits %>%
  mutate(time_to_next = as.numeric(difftime(lead(start), end, units = "mins")))
between_stats <- visits %>%
  summarise(
    min_time_to_next = min(time_to_next, na.rm = TRUE),
    max_time_to_next = max(time_to_next, na.rm = TRUE),
    avg_time_to_next = mean(time_to_next, na.rm = TRUE),
    sdev_time_to_next = sd(time_to_next, na.rm = TRUE))
```