---
# title: "Dineout EDA"
author: "Dillon"
date: "2023-11-20"
output: pdf_document
---

```{r setup, include=FALSE}
library(sf)
library(dplyr)
restaurants = read.csv("restaurants.csv")
movements = read.csv("movements.csv")

# buffers
buffer_full = 0.015
buffer_limited = 0.01
buffer_market = 0.05
buffer_supermarc = 0.05
buffer_warehouse = 0.05
buffer_conv = 0.01
```

# Categories of restaurants
```{r, echo=TRUE}
unique(restaurants$Category)
```

# Convert locations df with long. and lat. data into sf data object and add buffers around restaurants

```{r, echo=TRUE}
rest_sf <- st_as_sf(restaurants[, c("Name", "Category", "Longitude", "Latitude")], coords = c("Longitude", "Latitude"))
buffer = 0.002
rest_with_buffs <- st_buffer(rest_sf, dist = buffer)
plot(select(rest_with_buffs, -Category), col = "lightblue", main = "Restaurants")
#text(st_coordinates(rest_with_buffs)[, "X"], st_coordinates(rest_with_buffs)[, "Y"], labels = rest_with_buffs$Name, col = "red")
```
# Find points in movements that are inside restaurant's buffers
```{r, echo=TRUE}
movements_sf <- st_as_sf(movements, coords = c("longitude", "latitude"))
tmp <- st_within(movements_sf, rest_with_buffs, sparse = FALSE)
intersects <- as.data.frame(tmp)
```

# Reduce to only contain id of restaurant within (or 0 if not within any buffer)

```{r, echo=TRUE}
intersects_reduced <- max.col(intersects != 0, ties.method = "first")
intersects_reduced[!rowSums(intersects != 0) > 0] <- 0
intersects_reduced <- data.frame(location = intersects_reduced)
```

# View visits

```{r, echo=TRUE}
intersects_reduced[100, ]
intersects[100, ]

intersects_reduced[1, ]
intersects[1, ]

intersects_reduced <- cbind(intersects_reduced, movements$datetime)
head(intersects_reduced)

intersects_reduced
```

# Loop through data frame, start tmp visit variable when 'visit' starts. If long enough, continue. Consider short breaks (going out to car? etc) and excessive visits (job?)

```{r, echo=TRUE}
```